\chapter{Input and Output}

\input{reader-and-printer}
\input{format}
\input{stream}

\section{Extensions}
\label{sec:extensions}

\begin{macro}{with-input-from-file}{stream-name file-name \rest args
    \key direction \akeys \body body}[cl-ext]
  Evaluate BODY with STREAM-NAME bound to an input-stream from file
FILE-NAME. ARGS is passed directly to open.
\end{macro}

\begin{macro}{with-output-to-file}{stream-name file-name \rest args
    \key direction \akeys \body body}[cl-ext]
  Evaluate BODY with STREAM-NAME to an output stream on the file
FILE-NAME. ARGS is sent as is to the call te open. It will supersed
file-name if it exists.
\begin{devnote}
  The supersed thing is not originaly in alexandria, but I think it's
  more useful.
\end{devnote}
\end{macro}

\begin{function}{write-string-into-file}{string pathname \key if-exists if-does-not-exist external-format}[cl-ext]
  Write STRING to PATHNAME.

The EXTERNAL-FORMAT parameter will be passed to
ENCODING-KEYWORD-TO-NATIVE, see ENCODING-KEYWORD-TO-NATIVE to
possible values.
\end{function}

\begin{function}{read-file-into-string}{pathname \key buffer-size external-format}[cl-ext]
  Return the contents of PATHNAME as a fresh string.

The file specified by PATHNAME will be read one ELEMENT-TYPE
element at a time, the EXTERNAL-FORMAT and ELEMENT-TYPEs must be
compatible.

The EXTERNAL-FORMAT parameter will be passed to
ENCODING-KEYWORD-TO-NATIVE, see ENCODING-KEYWORD-TO-NATIVE to
possible values.
\end{function}

\begin{function}{copy-file}{from to \key if-to-exists element-type}[cl-ext]
  
\end{function}

\begin{function}{copy-stream}{input output \op element-type}[cl-ext]
  Reads data from FROM and writes it to TO. Both FROM and TO must be streams,
they will be passed to read-sequence/write-sequence and must have compatable
element-types.
\end{function}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
