\section{Streams}

\begin{function}{broadcast-stream-streams}{instance}
  
\end{function}

\begin{function}{clear-input}{\op stream}
  
\end{function}

\begin{function}{clear-output}{\op stream}
  
\end{function}

\begin{function}{close}{stream \key abort}
  Close the given STREAM. No more I/O may be performed, but
  inquiries may still be made. If :ABORT is true, an attempt is made
  to clean up the side effects of having created the stream.
\end{function}

\begin{function}{concatenated-stream-streams}{instance}
  
\end{function}

\begin{function}{echo-stream-input-stream}{instance}
  
\end{function}

\begin{function}{echo-stream-output-stream}{instance}
  
\end{function}

\begin{function}{file-length}{stream}
  
\end{function}

\begin{function}{file-position}{stream \op position}
  
\end{function}

\begin{function}{file-string-length}{stream object}
  
\end{function}

\begin{function}{finish-output}{\op stream}
  
\end{function}

\begin{function}{force-output}{\op stream}
  
\end{function}

\begin{function}{fresh-line}{\op stream}
  
\end{function}

\begin{function}{get-output-stream-string}{stream}
  
\end{function}

\begin{function}{input-stream-p}{stream}
  Can STREAM perform input operations?
\end{function}

\begin{function}{interactive-stream-p}{stream}
  Is STREAM an interactive stream?
\end{function}

\begin{function}{listen}{\op stream}
  
\end{function}

\begin{function}{make-broadcast-stream}{\rest streams}
  
\end{function}

\begin{function}{make-concatenated-stream}{\rest streams}
  Return a stream which takes its input from each of the streams in turn,
   going on to the next at EOF.
\end{function}

\begin{function}{make-echo-stream}{input-stream output-stream}
  Return a bidirectional stream which gets its input from INPUT-STREAM and
   sends its output to OUTPUT-STREAM. In addition, all input is echoed to
   the output stream.
\end{function}

\begin{function}{make-string-input-stream}{string \op start end}
  Return an input stream which will supply the characters of STRING between
  START and END in order.
\end{function}

\begin{function}{make-string-output-stream}{\key element-type \aux buffer}
  Return an output stream which will accumulate all output given it for the
benefit of the function GET-OUTPUT-STREAM-STRING.
\end{function}

\begin{function}{make-synonym-stream}{symbol}
  
\end{function}

\begin{function}{make-two-way-stream}{input-stream output-stream}
  Return a bidirectional stream which gets its input from INPUT-STREAM and
   sends its output to OUTPUT-STREAM.
\end{function}

\begin{function}{open}{filename \key direction element-type if-exists if-does-not-exist
 external-format \aux direction if-does-not-exist if-exists}
  Return a stream which reads from or writes to FILENAME.
  Defined keywords:
   :DIRECTION - one of :INPUT, :OUTPUT, :IO, or :PROBE
   :ELEMENT-TYPE - the type of object to read or write, default BASE-CHAR
   :IF-EXISTS - one of :ERROR, :NEW-VERSION, :RENAME, :RENAME-AND-DELETE,
                       :OVERWRITE, :APPEND, :SUPERSEDE or NIL
   :IF-DOES-NOT-EXIST - one of :ERROR, :CREATE or NIL
  See the manual for details.
\end{function}

\begin{function}{open-stream-p}{stream}
  Return true if STREAM is not closed. A default method is provided
  by class FUNDAMENTAL-STREAM which returns true if CLOSE has not been
  called on the stream.
\end{function}

\begin{function}{output-stream-p}{stream}
  Can STREAM perform output operations?
\end{function}

\begin{function}{peek-char}{\op peek-type stream eof-error-p eof-value recursive-p}
  
\end{function}

\begin{function}{read-byte}{stream \op eof-error-p eof-value}
  
\end{function}

\begin{function}{read-char}{\op stream eof-error-p eof-value recursive-p}
  
\end{function}

\begin{function}{read-char-no-hang}{\op stream eof-error-p eof-value recursive-p}
  
\end{function}

\begin{function}{read-line}{\op stream eof-error-p eof-value recursive-p}
  
\end{function}

\begin{function}{read-sequence}{seq stream \key start end}
  Destructively modify SEQ by reading elements from STREAM.
  That part of SEQ bounded by START and END is destructively modified by
  copying successive elements into it from STREAM. If the end of file
  for STREAM is reached before copying all elements of the subsequence,
  then the extra elements near the end of sequence are not updated, and
  the index of the next element is returned.
\end{function}

\begin{function}{stream-element-type}{stream}
  Return a type specifier for the kind of object returned by the
  STREAM. The class FUNDAMENTAL-CHARACTER-STREAM provides a default method
  which returns CHARACTER.
\end{function}

\begin{function}{stream-error-stream}{condition}
  
\end{function}

\begin{function}{stream-external-format}{stream}
  
\end{function}

\begin{function}{streamp}{stream}
  
\end{function}

\begin{function}{synonym-stream-symbol}{instance}
  
\end{function}

\begin{function}{terpri}{\op stream}
  
\end{function}

\begin{function}{two-way-stream-input-stream}{instance}
  
\end{function}

\begin{function}{two-way-stream-output-stream}{instance}
  
\end{function}

\begin{function}{unread-char}{character \op stream}
  
\end{function}

\begin{function}{write-byte}{integer stream}
  
\end{function}

\begin{function}{write-char}{character \op stream}
  
\end{function}

\begin{function}{write-line}{string \op stream \key start end}
  
\end{function}

\begin{function}{write-sequence}{seq stream \key start end}
  Write the elements of SEQ bounded by START and END to STREAM.
\end{function}

\begin{function}{write-string}{string \op stream \key start end}
  
\end{function}

\begin{function}{y-or-n-p}{\op format-string \rest arguments}
  Y-OR-N-P prints the message, if any, and reads characters from
   *QUERY-IO* until the user enters y or Y as an affirmative, or either
   n or N as a negative answer. It asks again if you enter any other
   characters.
\end{function}

\begin{function}{yes-or-no-p}{\op format-string \rest arguments}
  YES-OR-NO-P is similar to Y-OR-N-P, except that it clears the
   input buffer, beeps, and uses READ-LINE to get the strings
   YES or NO.
\end{function}

\begin{macro}{with-input-from-string}{var \body forms-decls}
  
\end{macro}

\begin{macro}{with-open-file}{stream \body body}
  
\end{macro}

\begin{macro}{with-open-stream}{var \body forms-decls}
  
\end{macro}

\begin{macro}{with-output-to-string}{var \body forms-decls}
  
\end{macro}

\begin{class}{broadcast-stream}{}
  
\end{class}

\begin{class}{concatenated-stream}{}
  
\end{class}

\begin{class}{echo-stream}{}
  
\end{class}

\begin{class}{file-stream}{}
  
\end{class}

\begin{class}{stream}{}
  
\end{class}

\begin{class}{string-stream}{}
  
\end{class}

\begin{class}{synonym-stream}{}
  
\end{class}

\begin{class}{two-way-stream}{}
  
\end{class}

\begin{variable}{*debug-io*}{}
  
\end{variable}

\begin{variable}{*error-output*}{}
  
\end{variable}

\begin{variable}{*query-io*}{}
  
\end{variable}

\begin{variable}{*standard-input*}{}
  
\end{variable}

\begin{variable}{*standard-output*}{}
  
\end{variable}

\begin{variable}{*terminal-io*}{}
  
\end{variable}

\begin{variable}{*trace-output*}{}
  
\end{variable}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: cl-dist-manual.tex
%%% End:
